---
title: "UreteroOutcomes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load data and packages, warning=FALSE}
rm(list=ls(all=TRUE))
library(dplyr)
library(lubridate)
library(ggplot2)
library(networkD3)

u <- data.table::fread("C:/Users/Courtney/Desktop/UreteroOutcomes/UreteroOutcomes.csv")
u$record_id <- as.character(u$record_id)

# Exclude people with insufficient records
exclude <- c("27","32","36","37","38","47","58","62","63","68","69","70","73","75","76","90","95","110","111","118","129","132","134","136","140","141","142","145","146","147","148")
length(exclude)
u <- filter(u,!(record_id %in% exclude))

# Label (main) one row per participant
u <- u %>% mutate(MainRow=ifelse(!is.na(sex),"Y","N"))

# Add columns for was prenatally diagnosed (to all rows)
pnppl <- unique((u %>% filter(diagnosis_reason___1 == 1))$record_id)
u <- u %>% mutate(PrenatDiagnosis=ifelse(record_id %in% pnppl,"Y","N"))
length(pnppl) == nrow(filter(u,MainRow=="Y"& is.na(date_diagnosis))) # make sure two ways of defining it give the same number

# Add column for ever had at least one surgery (to all rows)
surgppl <- unique((u %>% filter(!is.na(surgery_date)))$record_id)
u <- u %>% mutate(EverSurg=ifelse(record_id %in% surgppl,"Y","N"))

# Add column for surgery number (for each row that was a surgery)
surg_num <- filter(u,!is.na(surgery_date)) %>% arrange(record_id,surgery_date) %>% select(record_id,surgery_date)
surg_num$surgery_number <- ave(surg_num$record_id, surg_num$record_id, FUN=seq_along)
u <- left_join(u,surg_num,by=c("record_id","surgery_date"))

# Add column for that person's date of first surgery or NA if never had (to all rows)
firstsurg <- distinct(filter(u,surgery_number=="1") %>% select(record_id,surgery_date)) %>%
              rename(first_surgery_date=surgery_date)
u <- left_join(u,firstsurg,by=c("record_id"))

# Add column for preop reflux (to all rows)
rfppl <- unique(filter(u,EverSurg=="Y" & follow_up_date<first_surgery_date) %>%
        filter(vur_fu>0|vur_fu_2>0|initial_reflux_status>0|duplex_reflux>0|contra_reflux_grade>0))$record_id
u <- mutate(u,prereflux=ifelse(record_id %in% rfppl,"Y",ifelse(EverSurg=="Y","N","NA")))

# Add column for postop reflux (to all rows)
rfppl <- unique(filter(u,EverSurg=="Y" & follow_up_date>first_surgery_date) %>%
            mutate(temp=first_surgery_date+90) %>% filter(follow_up_date<=temp) %>%
        filter(vur_fu>0|vur_fu_2>0|initial_reflux_status>0|duplex_reflux>0|contra_reflux_grade>0))$record_id
u <- mutate(u,postreflux=ifelse(record_id %in% rfppl,"Y",ifelse(EverSurg=="Y","N","NA")))

# Add column for time from diagnosis (birth date if prenatdiag) to first surg (to Main row only)
dtfs <- filter(u,MainRow=="Y") %>% 
  mutate(daytofirstsurg=ifelse(PrenatDiagnosis=="Y",ymd(first_surgery_date)-ymd(dob),ymd(first_surgery_date)-ymd(date_diagnosis))) %>% select(record_id,daytofirstsurg)
u <- left_join(u,dtfs,by=c("record_id"))

# Add column for surg 1 type 1 and for those people, if it was definitive, to all columns
ppl <- (u %>% filter(surgery_number==1&type_surgery==1))$record_id
notdef <-  filter(u,record_id %in% ppl&surgery_number==2)$record_id
u$surg1type1 <- ifelse(u$record_id %in% ppl,"Y","N")
u$TUIdef <- ifelse(u$record_id %in% ppl & u$record_id %in% notdef,"N",ifelse(u$record_id %in% ppl,"Y","NA"))
u$TUIdefnum <- as.numeric(ifelse(u$record_id %in% ppl & u$record_id %in% notdef,"0",ifelse(u$record_id %in% ppl,"1","NA")))

# Add column for add at first surg to main row
as <- filter(u,MainRow=="Y") %>% mutate(agesurg1=as.numeric(ymd(first_surgery_date)-ymd(dob))/365.25)%>% select(record_id,agesurg1)
u <- left_join(u,as,by=c("record_id"))

```


```{r primary TUI surgery dates}
temp <- filter(u,surg1type1=="Y"&MainRow=="Y")
ggplot(temp,aes(ymd(first_surgery_date)))+geom_histogram(color="black",binwidth=365.25)+ # binned to bars of 1 year widths
  labs(x="Date of Primary TUI",y="Count")+theme_classic()
ggsave("date_primaryTUI_hist.png")

```

```{r Demographics of surg but not primary TUI people}

nonTUIppl <- filter(u,surg1type1=="N"&surgery_number=="1")$record_id

# Breakdown by sex and prenat diagnosis
filter(u,record_id %in% nonTUIppl & MainRow=="Y") %>% group_by(PrenatDiagnosis) %>% count()
filter(u,record_id %in% nonTUIppl & MainRow=="Y") %>% group_by(sex) %>% count()

# Time to diagnosis
temp <- filter(u,record_id %in% nonTUIppl &PrenatDiagnosis=="N"&MainRow=="Y")
median((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age) # days
mean((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)
min((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)
max((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)

# UTI
ever_uti_nonTUI <- unique(filter(u, record_id %in% nonTUIppl & (diagnosis_reason___2=="1" | reason_imm_surgery___4=="1" | uti=="1" | febrile_uti=="1" | post_op_uti=="1" | repeat_surg_indication=="1" | reason_surgery___4 == "1"))$record_id)
length(unique(ever_uti_nonTUI))

# Age at first surg for non primary TUI
median(filter(u,surg1type1=="N"&surgery_number=="1")$agesurg1)
min(filter(u,surg1type1=="N"&surgery_number=="1")$agesurg1)
max(filter(u,surg1type1=="N"&surgery_number=="1")$agesurg1)
mean(filter(u,surg1type1=="N"&surgery_number=="1")$agesurg1)

```

```{r Demongraphics of nonsurg people}

# Breakdown by sex and prenat diagnosis
filter(u,EverSurg=="N"&MainRow=="Y") %>% group_by(PrenatDiagnosis) %>% count()
filter(u,EverSurg=="N"&MainRow=="Y") %>% group_by(sex) %>% count()

# Time to diagnosis
temp <- filter(u,EverSurg=="N"&PrenatDiagnosis=="N"&MainRow=="Y")
median((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age) # days
mean((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)
min((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)
max((temp %>% mutate(diag_age=as.numeric(ymd(date_diagnosis)-ymd(dob))))$diag_age)

# UTI
non_surg_ppl <- filter(u,EverSurg=="N"&MainRow=="Y")$record_id
ever_uti_nonsurg <- unique(filter(u, record_id %in% non_surg_ppl & (diagnosis_reason___2=="1" | uti=="1" | febrile_uti=="1"))$record_id)

```


```{r Med Summary Stats}
# How many were prenatally diagnosed?
nrow(filter(u,MainRow=="Y") %>% filter(PrenatDiagnosis=="Y"))

# How many had at least one surgery performed?
nrow(filter(u,MainRow=="Y") %>% filter(EverSurg=="Y"))

# What sides of ureteroceles?
filter(u,MainRow=="Y") %>% group_by(kidney_with_ureterocele) %>% count()

# What types of ureteroceles did ppl have?
filter(u,MainRow=="Y") %>% group_by(ureterocele_pole) %>% count()
filter(u,MainRow=="Y") %>% group_by(ureterocele_site) %>% count()
u_types <- filter(u,MainRow=="Y") %>% group_by(kidney_with_ureterocele,ureterocele_pole,ureterocele_site) %>% count()
write.table(u_types,"C:/Users/Courtney/Desktop/UreteroOutcomes/ureterotypes.tsv", quote=F, sep="\t", row.names=F, col.names=T)

# How many who had surgery needed additional surgery
length(unique(filter(u,surgery_number=="2")$record_id)) # had 2 or more
length(unique(filter(u,surgery_number=="3")$record_id)) # had 3

# How many unique individuals had UTI directly after surgery at least once?
length(unique((u %>% filter(post_op_uti=="1"))$record_id))

# Average time of followup for all
avgfu <- filter(u,PrenatDiagnosis=="Y"&MainRow=="Y") %>% select(record_id,dob,date_visit_4)
avgfu <- avgfu %>% mutate(fu=ymd(date_visit_4)-ymd(dob)) %>% select(record_id,fu)
temp <- filter(u,PrenatDiagnosis=="N") %>% select(record_id,follow_up_date) %>% arrange(follow_up_date) %>% group_by(record_id) %>% filter(1:n() == 1)

# Avg time followup for type 1 surg 1 after first surg
ppl <- (u %>% filter(surgery_number==1&type_surgery==1))$record_id
temp <-  filter(u,record_id %in% ppl & MainRow=="Y") %>% select(record_id,first_surgery_date,date_visit_4)
temp <- temp %>% mutate(fu=ymd(date_visit_4)-ymd(first_surgery_date))
temp %>% summarize(meanfu=mean(fu),medfu=median(fu),sdfu=sd(fu),rangefu=range(fu))

# How many type 1 surg 1 had post op uti for that  & other than those how many had a febrile uti at some point after
pou <- unique((u %>% filter(surgery_number==1&type_surgery==1  & post_op_uti=="1"))$record_id)
length(pou)
ppl <- (u %>% filter(surgery_number==1&type_surgery==1))$record_id
length(unique(filter(u,record_id %in% ppl & follow_up_date>first_surgery_date & febrile_uti==1 & !(record_id %in% pou))$record_id))

########### SURG 1 TYPE 1

# Age at first surg for type 1 surg1
s <- filter(u,MainRow=="Y"&surg1type1=="Y")
median(s$agesurg1)
mean(s$agesurg1)
sd(s$agesurg1)
min(s$agesurg1)
max(s$agesurg1)

# Fraction prenatally diagnosed type1 surg1
table(s %>% select(PrenatDiagnosis))

# How many diagnosed because of UTIs
table(s %>% select(diagnosis_reason___2))

# How many incidental diagnosed (but not prenat diagnosed)
nrow(s %>% filter(diagnosis_reason___4=="1"&diagnosis_reason___1=="0"))

# Avg age diagnosis for non prenatally diagnosed type1 surg1
a <- filter(s,PrenatDiagnosis=="N") %>% select(record_id,dob,date_diagnosis)
a <- a %>% mutate(Age_Diagnosis=as.numeric((ymd(a$date_diagnosis)-ymd(a$dob))))
mean(na.omit(a$Age_Diagnosis))/365.25
median(na.omit(a$Age_Diagnosis))/365.25 
sd(na.omit(a$Age_Diagnosis))/365.25
min(na.omit(a$Age_Diagnosis))/365.25
max(na.omit(a$Age_Diagnosis))/365.25

# How many operated at first visit Surg 1 type 1
nrow(s %>% filter(treatment_plan=="1"))

# How many each side ureterocele surg1 type`
s %>% group_by(kidney_with_ureterocele) %>% count()

# Avg num of followups
ppl <- (u %>% filter(surgery_number==1&type_surgery==1))$record_id
n <- (filter(u,record_id %in% ppl) %>% group_by(record_id) %>% count())$n
summary(n)

```

```{r Surgery type}

# For each surgery number (first surgery, second, or third surgery) what was the breakdown of each surgery type?
surg_num_bkd <- u %>% group_by(surgery_number,type_surgery) %>% count()
surg_num_bkd
write.table(surg_num_bkd,"C:/Users/Courtney/Desktop/UreteroOutcomes/surgtypebreakdown.tsv", quote=F, sep="\t", row.names=F, col.names=T)

# How many people had surgery type 1 on first surgery then had a second surgery
ppl <- unique(filter(u,surgery_number=="1"&type_surgery=="1")$record_id)
filter(u,record_id %in% ppl & surgery_number=="2")

# What was the surgery info (for all of their surgeries) for people who had 3 surgeries
ppl <- unique(filter(u,surgery_number=="3")$record_id)
three_surg <- filter(u,record_id %in% ppl & !is.na(surgery_date)) %>% select(record_id,surgery_number,type_surgery,surgery_date)
three_surg
write.table(three_surg,"C:/Users/Courtney/Desktop/UreteroOutcomes/threesurgtypebreakdown.tsv", quote=F, sep="\t", row.names=F, col.names=T)

# Is there an association with the surgery type of a first surgery (had type 1 vs not type 1) and the need to have additional?
nottype1 <- data.frame(record_id=unique(filter(u,surgery_number=="1"&type_surgery!="1")$record_id))
type1 <- data.frame(record_id=unique(filter(u,surgery_number=="1"&type_surgery=="1")$record_id))
nottype1hadmultisurg <- right_join(u,nottype1,by=c("record_id")) %>% filter(surgery_number=="2")
type1hadmultisurg <- right_join(u,type1,by=c("record_id")) %>% filter(surgery_number=="2")

df <- data.frame("needed extra surg" = c(nrow(type1hadmultisurg), nrow(nottype1hadmultisurg)), "didn't need extra surg" = c(nrow(type1) - nrow(type1hadmultisurg), nrow(nottype1) - nrow(nottype1hadmultisurg)), row.names = c("surg num1 type 1", "surg num1 type not 1"))
df
fisher.test(df,alternative="greater") # was this enriched relative to expected

```

```{r UTI timing relative to surg}
# Of the patients who had a UTI and underwent a surgery, how many had their first UTI before surgery and how many after?
surg_ppl <- unique(filter(u,EverSurg=="Y")$record_id)

# BEFORE
tempsurg <- u %>% filter(!is.na(surgery_date)) %>% arrange(record_id,surgery_date) %>% group_by(record_id) %>% filter(row_number()==1) %>% select(record_id,surgery_date)  # earliest surg
tempfu <- u %>% filter(febrile_uti=="1"|uti=="1") %>% arrange(record_id,follow_up_date) %>% group_by(record_id) %>% filter(row_number()==1) %>% select(record_id,follow_up_date)  # earliest followup w uti
temp <- right_join(tempsurg,tempfu,by=c("record_id"))
crit3ppl <- unique(filter(temp,follow_up_date<surgery_date)$record_id)
before <- filter(u, record_id %in% surg_ppl & (diagnosis_reason___2=="1" | reason_imm_surgery___4=="1" | record_id %in% crit3ppl))
length(unique(before$record_id)) # num people before

# AFTER
ever_uti_surg <- unique(filter(u, record_id %in% surg_ppl & (diagnosis_reason___2=="1" | reason_imm_surgery___4=="1" | uti=="1" | febrile_uti=="1" | post_op_uti=="1" | repeat_surg_indication=="1" | reason_surgery___4 == "1"))$record_id)
aftersurg <- unique(setdiff(ever_uti_surg,unique(before$record_id)))
length(aftersurg) # had uti after surg

u_full <- u %>% mutate(firstUTIbeforesurg=ifelse(record_id %in% unique(before$record_id),"Y","N"))
u_full <- u_full %>% mutate(firstUTIaftersurg=ifelse(record_id %in% aftersurg,"Y","N"))
u_full <- u_full %>% mutate(EverUTI=ifelse(record_id %in% ever_uti_surg,"Y","N"))

```

```{r Age and Type of Diagnosis for non-PrenatDiagnosis}
# Average age at diagnosis if not prenatally diagnosed
a <- filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(record_id,dob,date_diagnosis)
a <- a %>% mutate(Age_Diagnosis=as.numeric((ymd(a$date_diagnosis)-ymd(a$dob))))
mean(na.omit(a$Age_Diagnosis)) # 412 days
median(na.omit(a$Age_Diagnosis)) # 68 days
ggplot(a,aes(x=Age_Diagnosis)) + geom_histogram(color="black", fill="white",binwidth=15) +
  geom_vline(aes(xintercept=mean(Age_Diagnosis)),
             color="blue", linetype="dashed", size=1)+
 geom_vline(aes(xintercept=median(Age_Diagnosis)),
          color="red", linetype="dashed", size=1)+
  labs(x="Average Age of Diagnosis (Days)",y="Number") + theme_classic()
ggsave("C:/Users/Courtney/Desktop/UreteroOutcomes/age_diagnosis15.png")

## How were people who were not prenatally diagnosed most frequently diagnosed?
# num individuals diagnosis_reason___2: 13 total, and of these 13 two are just diagnosis_reason___2 and diagnosis_reason___4, and one is diagnosis_reason___2 and diagnosis_reason___4 and diagnosis_reason___99
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___2=="1")
# num individuals diagnosis_reason___3: 2 total, one of these is diagnosis_reason___3 and diagnosis_reason___4 and one of these is diagnosis_reason___3 and diagnosis_reason___99
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___3=="1")
# num individuals only diagnosis_reason___4: 43 total, of these 18 are also diagnosis_reason___99
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___4=="1"&diagnosis_reason___1=="0"&diagnosis_reason___2=="0"&diagnosis_reason___3=="0")
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___4=="1"&diagnosis_reason___99=="1"&diagnosis_reason___1=="0"&diagnosis_reason___2=="0"&diagnosis_reason___3=="0")
# num individuals with diagnosis_reason___98: 0
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___98=="1")
# num individuals just diagnosis_reason___99: 3
filter(u,MainRow=="Y"&PrenatDiagnosis=="N") %>% select(matches("diagnosis_reason")) %>% filter(diagnosis_reason___4=="0"&diagnosis_reason___1=="0"&diagnosis_reason___2=="0"&diagnosis_reason___3=="0"&diagnosis_reason___99=="1")

```

```{r Overall UTI}

# What fraction of people that had a UTI directly after their first surgery then went on to have at least one more: 1/2
multsurg <- filter(u,surgery_number==2)$record_id
length(unique((u %>% filter(post_op_uti=="1",record_id %in% multsurg))$record_id))/length(unique((u %>% filter(post_op_uti=="1"))$record_id))
matrix(c(4,8-4,26-4,86-8), nrow=2)
fisher.test(matrix(c(4,8-4,26-4,86-8), nrow=2)) # was this enriched relative to expected; odds ratio 3.5, P-value 0.09

# How many people who had at least one surgery ever had a UTI and is this enriched relative to those who never had surgery
surg <- unique((filter(u,EverSurg=="Y")$record_id))
nosurg <- setdiff(unique((filter(u,is.na(surgery_date))$record_id)),surg)
surg_uti <- filter(u,record_id %in% surg & (post_op_uti=="1"|febrile_uti=="1"|uti=="1")) %>% select(record_id,surgery_date,matches("uti"))
length(unique(surg_uti$record_id)) # 47
nosurg_uti <- filter(u,record_id %in% nosurg & (post_op_uti=="1"|febrile_uti=="1"|uti=="1")) %>% select(record_id,surgery_date,matches("uti"))
length(unique(nosurg_uti$record_id)) # 7
matrix(c(47,86-47,47+7-47,123-86), nrow=2)
fisher.test(matrix(c(47,86-47,47+7-47,123-86), nrow=2)) # was this enriched relative to expected; 6.3 odds ratio, P-value 2.5e-5

# What is the average number of UTIs (total and by surg or nosurg)
surg_uti_stat <- as.data.frame(filter(u,record_id %in% surg) %>% group_by(record_id) %>% count())
surg_uti_stat %>% summarize(mean(n),median(n),sd(n)) # 10.9,10,5.3
nosurg_uti_stat <- as.data.frame(filter(u,record_id %in% nosurg) %>% group_by(record_id) %>% count())
nosurg_uti_stat %>% summarize(mean_n=mean(n),median_n=median(n),sd_n=sd(n)) # 6.2,5,2.8
uti_stat <- bind_rows(surg_uti_stat %>% mutate(Surg=as.factor("Yes")),nosurg_uti_stat %>% mutate(Surg=as.factor("No")))
ggplot(uti_stat,aes(Surg,n))+geom_violin(trim=FALSE,aes(fill=Surg))+
  labs(x="Individual Ever Had Surgery?",y="Number of UTIs")+
  scale_fill_manual(values=c("purple", "#E69F00"))+geom_boxplot(width=0.1)+theme_classic()
ggsave("C:/Users/Courtney/Desktop/UreteroOutcomes/uti_bysurg.png")
```

```{r Prenat Diagnosis Surg Outcomes}

# Is a prenatal diagnosis more likely than diagnosis at a later date: undergoing at least one surgery
notprenat <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="N")$record_id))
prenat <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="Y")$record_id))
notprenathadsurg <- unique((right_join(u,notprenat,by=c("record_id")) %>% filter(EverSurg=="Y"))$record_id)
prenathadsurg <- unique((right_join(u,prenat,by=c("record_id")) %>% filter(EverSurg=="Y"))$record_id)

df <- data.frame("Had at least one surgery" = c(length(prenathadsurg), length(notprenathadsurg)), "Didn't need surg" = c(nrow(prenat) - length(prenathadsurg), nrow(notprenat) - length(notprenathadsurg)), row.names = c("Prenatally Diagnosed", "Not Prenatally Diagnosed"))
df
fisher.test(df) # was this enriched relative to expected

# For patients who underwent surgery, was a prenatal diagnosis more likely than diagnosis at a later date to need more than one surgery?
notprenatsurg <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="N"&EverSurg=="Y")$record_id))
prenatsurg <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="Y"&EverSurg=="Y")$record_id))
notprenatmultisurg <- unique((right_join(u,notprenatsurg,by=c("record_id")) %>% filter(surgery_number=="2"))$record_id)
prenatmultisurg <- unique((right_join(u,prenatsurg,by=c("record_id")) %>% filter(surgery_number=="2"))$record_id)

df <- data.frame("Had multi surgery" = c(length(prenatmultisurg), length(notprenatmultisurg)), "Only one surg" = c(nrow(prenatsurg) - length(prenatmultisurg), nrow(notprenatsurg) - length(notprenatmultisurg)), row.names = c("Prenatally Diagnosed", "Not Prenatally Diagnosed"))
df
fisher.test(df) # was this enriched relative to expected

# For primary tui patients who underwent surgery, was a prenatal diagnosis more likely than diagnosis at a later date to need more than one surgery?
notprenattui <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="N"&surg1type1=="Y")$record_id))
prenattui <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="Y"&surg1type1=="Y")$record_id))
notprenatmultisurg <- unique((right_join(u,notprenattui,by=c("record_id")) %>% filter(surgery_number=="2"))$record_id)
prenatmultisurg <- unique((right_join(u,prenattui,by=c("record_id")) %>% filter(surgery_number=="2"))$record_id)

df <- data.frame("Had multi surgery" = c(length(prenatmultisurg), length(notprenatmultisurg)), "Only primary TUI" = c(nrow(prenattui) - length(prenatmultisurg), nrow(notprenattui) - length(notprenatmultisurg)), row.names = c("Prenatally Diagnosed", "Not Prenatally Diagnosed"))
df
fisher.test(df) # was this enriched relative to expected

```

```{r Time Diagnosis to Surg Outcomes}
# Note: for the below, set diagnosis date to date of birth for prenat diagnosed

# Is a greater time difference between diagnosis and first surgery significantly related to undergoing additional surgeries?
multisurg <- unique(filter(u,surgery_number=="2")$record_id)
temp <- u %>% filter(MainRow=="Y"&EverSurg=="Y") %>% mutate(MultiSurg=ifelse(record_id %in% multisurg,"Yes","No"))
ggplot(temp,aes(x=MultiSurg,y=daytofirstsurg))+geom_violin(trim=FALSE,aes(fill=MultiSurg))+
  labs(x="Individual Had Multiple Surgeries?",y="Days To First Surgery From Diagnosis")+
  scale_fill_manual(values=c("purple", "#E69F00", "blue"))+geom_boxplot(width=0.1)+theme_classic()
ggsave("C:/Users/Courtney/Desktop/UreteroOutcomes/multisurg_timetosurg.png")
t.test(filter(temp,MultiSurg=="Yes")$daytofirstsurg, filter(temp,MultiSurg=="No")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway

# Is a greater time difference between diagnosis and first surgery significantly related to UTI after first surg
temp <- filter(u_full,MainRow=="Y")
ggplot(temp,aes(x=firstUTIaftersurg,y=daytofirstsurg))+geom_violin(trim=FALSE,aes(fill=firstUTIaftersurg))+
  labs(x="Individual Had Any UTI after First Surgery?",y="Days To First Surgery From Diagnosis")+
  scale_fill_manual(values=c("purple", "#E69F00", "blue"))+geom_boxplot(width=0.1)+theme_classic()
ggsave("C:/Users/Courtney/Desktop/UreteroOutcomes/uti_timetosurg.png")
t.test(filter(temp,firstUTIaftersurg=="Y")$daytofirstsurg, filter(temp,firstUTIaftersurg=="N")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway


```

```{r Reflux Outcomes}
newrefluxppl <- filter(u,MainRow=="Y"&prereflux=="N"&postreflux=="Y")$record_id

# For patients who underwent surgery, was a prenatal diagnosis more likely than diagnosis at a later date to develop NEW reflux within 3 months after surgery
notprenat <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="N")$record_id))
prenat <- data.frame(record_id=unique(filter(u,PrenatDiagnosis=="Y")$record_id))
notprenatreflux <- unique((right_join(u,notprenat,by=c("record_id")) %>% filter(record_id %in% newrefluxppl))$record_id)
prenatreflux <- unique((right_join(u,prenat,by=c("record_id")) %>% filter(record_id %in% newrefluxppl))$record_id)

df <- data.frame("Had New Reflux" = c(length(prenatreflux), length(notprenatreflux)), "Didn't Have New Reflux" = c(nrow(prenat) - length(prenatreflux), nrow(notprenat) - length(notprenatreflux)), row.names = c("Prenatally Diagnosed", "Not Prenatally Diagnosed"))
df
fisher.test(df) # was this enriched relative to expected

# Is a greater time difference between diagnosis and first surgery significantly related to developing NEW reflux within 3 months after surgery
temp <- filter(u,MainRow=="Y") %>% mutate(Reflux=paste0("pre:",prereflux,"_post:",postreflux))
ggplot(temp,aes(x=Reflux,y=daytofirstsurg))+geom_violin(trim=FALSE,aes(fill=Reflux))+
  labs(x="Reflux?",y="Days To First Surgery From Diagnosis")+
  scale_fill_manual(values=c("purple", "#E69F00", "blue"))+geom_boxplot(width=0.1)+theme_classic()
ggsave("C:/Users/Courtney/Desktop/UreteroOutcomes/reflux_timetosurg.png")

t.test(filter(temp,Reflux=="pre:N_post:Y")$daytofirstsurg, filter(temp,Reflux=="pre:N_post:N")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway

t.test(filter(temp,Reflux=="pre:N_post:Y")$daytofirstsurg, filter(temp,Reflux!="pre:N_post:Y")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway

t.test(filter(temp,Reflux=="pre:N_post:N")$daytofirstsurg, filter(temp,Reflux!="pre:N_post:N")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway

t.test(filter(temp,prereflux=="N")$daytofirstsurg, filter(temp,prereflux=="Y")$daytofirstsurg, var.equal = TRUE) # need to confirm assumptions for this t test anyway
```

```{r Antibiotics}
one <- unique(filter(u,antibiotics=="1")$record_id)
zero <- unique(filter(u,antibiotics=="0")$record_id)
zero <- setdiff(zero,one)

t.test(filter(u,record_id %in% one & MainRow=="Y")$daytofirstsurg,filter(u,record_id %in% zero & MainRow=="Y")$daytofirstsurg)

onehadsurg <- filter(u,MainRow=="Y" & record_id %in% one & EverSurg=="Y")$record_id
zerohadsurg <- filter(u,MainRow=="Y" & record_id %in% zero & EverSurg=="Y")$record_id

df <- data.frame("Had at least one surgery" = c(length(zerohadsurg), length(onehadsurg)), "Didn't need surg" = c(length(zero) - length(zerohadsurg), length(one) - length(onehadsurg)), row.names = c("Antibiotics at least one 1", "Antibiotics only ever 0"))
df
fisher.test(df) # was this enriched relative to expected

# How many people have a followup within 90 days of first surgery
length(unique(filter(u,follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90)$record_id))

### For had type 1 as first surg people
# How many people have a followup within 90 days of first surgery
type1firstsurgppl <- unique(filter(u,surgery_number=="1"&type_surgery=="1")$record_id)
length(unique(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90)$record_id))
# How many had more than 1 followup within 90 days 
filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90) %>% group_by(record_id) %>% count() %>% arrange(-n) %>% filter(n>1)

# did anyone with multiple followups within 90 days have dif antibiotic status in the followups (if no then can just filter to one visit per person and count from there for next step)
distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90) %>% select(record_id,antibiotics)) %>% group_by(record_id) %>% count() %>% filter(n>1)
# note: record_id==20 had one followup on antibiotics and one not so will count that as on because was on at least once within 90 days after first surg

# What frac of people were on antibiotics within 90 days of first surg
distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90) %>% select(record_id,antibiotics)) %>% filter(!(record_id=="20" & antibiotics=="0")) %>% group_by(antibiotics) %>% count() %>% filter(n>1)

# Who were the 4 not on antibiotics and was TUI definitive for them vs abx ppl
notabx<-(distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90) %>% select(record_id,antibiotics)) %>% filter(!(record_id=="20" & antibiotics=="0"))  %>% filter(antibiotics=="0"))$record_id
nrow(filter(u,record_id %in% notabx & surgery_number=="2"))
abx<-(distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+90) %>% select(record_id,antibiotics)) %>% filter(!(record_id=="20" & antibiotics=="0"))  %>% filter(antibiotics=="1"))$record_id
nrow(filter(u,record_id %in% abx & surgery_number=="2"))

### For had type 1 as first surg people
# How many people have a followup within 365 days of first surgery
numdays=365
type1firstsurgppl <- unique(filter(u,surgery_number=="1"&type_surgery=="1")$record_id)
length(unique(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays)$record_id))
# How many had more than 1 followup within 365 days 
filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% group_by(record_id) %>% count() %>% arrange(-n) %>% filter(n>1)

# did anyone with multiple followups within 365 days have dif antibiotic status in the followups (if no then can just filter to one visit per person and count from there for next step)
multi <- (distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% select(record_id,antibiotics)) %>% group_by(record_id) %>% count() %>% filter(n>1))$record_id
# if anyone has multiple then based on how it's coded at least one was on antibiotic, so count them as yes antibiotic

# What frac of people were on antibiotics within 365 days of first surg
distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% select(record_id,antibiotics)) %>% filter(!(record_id %in% multi & antibiotics=="0")) %>% group_by(antibiotics) %>% count() %>% filter(n>1)

# How many within first year of surg had an appointment where they werent on antibiotics
noan <- unique(((distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% select(record_id,antibiotics))) %>% filter(antibiotics=="0"))$record_id)

# How many people never had a followup where they were on antibiotics in the first year
anti <- setdiff(type1firstsurgppl,unique(((distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% select(record_id,antibiotics))) %>% filter(antibiotics=="1"))$record_id))

# How many had an appointment where they discontinued antibiotics within first year of surg
dis <- unique((distinct(filter(u,record_id %in% type1firstsurgppl & follow_up_date>first_surgery_date&follow_up_date<first_surgery_date+numdays) %>% select(record_id,antibiotics,antibiotics_end)) %>% filter(antibiotics_end=="3"))$record_id)

intersect(noan,anti) # 3 people had no followups in first year, 3 had at least one followup where they were not on antibiotics in first year and no followups where were on antibiotics in first year, 13 had at least one followup where were not on antibiotics in first year and at least one followup where was on antibiotics in first year

setdiff(dis,noan) # 5 people had an appointment where they discontinued antibiotics within first year of surg that didnt also have an appointment where they were off antibiotics (8 others also had appointment where discontinued them but they also had an appointment where were off antibiotics so they are already counted in that number)
```

```{r Prenat Diagnosis}

# Relationship between day to first surgery and prenatally diagnosed
ppl1 <- unique(filter(u,PrenatDiagnosis=="Y")$record_id)
ppl2 <- unique(filter(u,PrenatDiagnosis=="N")$record_id)

t.test(filter(u,record_id %in% ppl1 & MainRow=="Y")$daytofirstsurg,filter(u,record_id %in% ppl2 & MainRow=="Y")$daytofirstsurg)

# Associations for prenatal diagnosis with ever having surgery
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & EverSurg=="Y")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & EverSurg=="Y")$record_id

df <- data.frame("Did X" = c(length(ppl2didx), length(ppl1didx)), "Didn't X" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("People 1", "People 2"))
df
fisher.test(df) # was this enriched relative to expected
```

```{r surgery type 1}
type1firstsurgppl <- unique(filter(u,surgery_number=="1"&type_surgery=="1")$record_id)

## Things associated with multiple surg for ppl who had surg type 1 on their 1st
ppl1 <- unique(filter(u,record_id %in% type1firstsurgppl & surgery_number=="2")$record_id)
ppl2 <- setdiff(type1firstsurgppl,ppl1)

## ureteropole 1
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & ureterocele_pole=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & ureterocele_pole=="1")$record_id

df <- data.frame("U pole 1" = c(length(ppl2didx), length(ppl1didx)), "Not U pole 1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df, alternative = "greater") # was this enriched relative to expected

## ureteropole site
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & ureterocele_site=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & ureterocele_site=="1")$record_id

df <- data.frame("U Site 1" = c(length(ppl2didx), length(ppl1didx)), "Not U Site 1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df, alternative = "greater") # was this enriched relative to expected

## prereflux
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & prereflux=="N")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & prereflux=="N")$record_id

df <- data.frame("Preop reflux == N" = c(length(ppl2didx), length(ppl1didx)), "Not Preop reflux == N" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

## proph_abx
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & proph_abx=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & proph_abx=="1")$record_id

df <- data.frame("proph_abx 1" = c(length(ppl2didx), length(ppl1didx)), "Not proph_abx 1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

## sex
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & sex=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & sex=="1")$record_id

df <- data.frame("sex 1" = c(length(ppl2didx), length(ppl1didx)), "Not sex 1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

## diagnosis_reason___1
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & diagnosis_reason___1=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & diagnosis_reason___1=="1")$record_id

df <- data.frame("diagnosis_reason___1 1" = c(length(ppl2didx), length(ppl1didx)), "Not diagnosis_reason___1 1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

## First UTI preop
ppl1didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl1 & firstUTIbeforesurg=="Y")$record_id
ppl2didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl2 & firstUTIbeforesurg=="Y")$record_id

df <- data.frame("firstUTIbeforesurg" = c(length(ppl2didx), length(ppl1didx)), "Not firstUTIbeforesurg" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

# EverUTI
ppl1didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl1 & EverUTI=="Y")$record_id
ppl2didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl2 & EverUTI=="Y")$record_id

df <- data.frame("EverUTI" = c(length(ppl2didx), length(ppl1didx)), "Not EverUTI" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df,alternative = "less") # was this enriched relative to expected

# First UTI postop
ppl1didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl1 & firstUTIaftersurg=="Y")$record_id
ppl2didx <- filter(u_full,MainRow=="Y" & record_id %in% ppl2 & firstUTIaftersurg=="Y")$record_id

df <- data.frame("firstUTIaftersurg" = c(length(ppl2didx), length(ppl1didx)), "Not firstUTIaftersurg" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df,alternative="less") # was this enriched relative to expected

# kidney_with_ureterocele
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & kidney_with_ureterocele=="3")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & kidney_with_ureterocele=="3")$record_id

df <- data.frame("kidney_with_ureterocele" = c(length(ppl2didx), length(ppl1didx)), "Not kidney_with_ureterocele" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

# treatment_plan
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & treatment_plan=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & treatment_plan=="1")$record_id

df <- data.frame("treatment_plan" = c(length(ppl2didx), length(ppl1didx)), "Not treatment_plan" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

# post_op_uti
ppl1didx <- filter(u,MainRow=="Y" & record_id %in% ppl1 & post_op_uti=="1")$record_id
ppl2didx <- filter(u,MainRow=="Y" & record_id %in% ppl2 & post_op_uti=="1")$record_id

df <- data.frame("post_op_uti" = c(length(ppl2didx), length(ppl1didx)), "Not post_op_uti" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

## Association with days until first surg with multi surg
t.test(filter(u,record_id %in% ppl1 & MainRow=="Y")$daytofirstsurg,filter(u,record_id %in% ppl2 & MainRow=="Y")$daytofirstsurg)



```
Technique contains the word "holmium" or "laser" or "fiber" (the two categories would be yes it contains at least one of those words, or no it contains words but not any of those three)
technique contains the word "electrode" or "bugbee" or "cautery" (same approach as above. specifically look at patients where technique has words in it and then it's YES if it contains one of those words or NO if it has words but not electrode or bugbee or cautery)
same as above but the words "needle" or "knife" or "can"

```{r technique}
techppl <- unique(filter(u,record_id %in% type1firstsurgppl & technique!="")$record_id)

## Things associated with multiple surg for ppl who had surg type 1 on their 1st
ppl1 <- unique(filter(u,record_id %in% techppl & surgery_number=="2")$record_id)
ppl2 <- setdiff(techppl,ppl1)

u_tech <- filter(u,record_id %in% techppl & surgery_number=="1")
laser <- unique(filter(u_tech,grepl("holmium|laser|fiber",technique))$record_id)
knife <- unique(filter(u_tech,grepl("needle|knife",technique))$record_id)
electrode <- unique(filter(u_tech,grepl("electrode|bugbee|cautery",technique))$record_id)

length(laser)
length(knife)
length(electrode)

temp <- filter(u,MainRow=="Y") %>% mutate(laser=ifelse(record_id %in% laser,"Y","N"))
temp <- temp %>% mutate(knife=ifelse(record_id %in% knife,"Y","N"))
temp <- temp %>% mutate(electrode=ifelse(record_id %in% electrode,"Y","N"))

# laser
ppl1didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl1 & laser=="Y")$record_id)
ppl2didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl2 & laser=="Y")$record_id)
lppl1didx <- ppl1didx
lppl2didx <- ppl2didx
length(ppl2didx)/(length(ppl2didx)+length(ppl1didx))

df <- data.frame("laser" = c(length(ppl2didx), length(ppl1didx)), "Not laser" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

# knife
ppl1didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl1 & knife=="Y")$record_id)
ppl2didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl2 & knife=="Y")$record_id)
length(ppl2didx)/(length(ppl2didx)+length(ppl1didx))

df <- data.frame("knife" = c(length(ppl2didx), length(ppl1didx)), "Not knife" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

# electrode
ppl1didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl1 & electrode=="Y")$record_id)
ppl2didx <- unique(filter(temp,MainRow=="Y" & record_id %in% ppl2 & electrode=="Y")$record_id)
eppl1didx <- ppl1didx
eppl2didx <- ppl2didx
length(ppl2didx)/(length(ppl2didx)+length(ppl1didx))

df <- data.frame("electrode" = c(length(ppl2didx), length(ppl1didx)), "Not electrode" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df, alternative="less") # was this enriched relative to expected

# Just compare electrode and laser ppl
df <- data.frame("electrode" = c(length(eppl2didx), length(eppl1didx)), "Laser" = c(length(lppl2didx), length(lppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected

```


```{r puncture}
## Things associated with multiple surg for ppl who had surg type 1 on their 1st
ppl1 <- unique(filter(u,record_id %in% type1firstsurgppl & surgery_number=="2")$record_id)
ppl2 <- setdiff(type1firstsurgppl,ppl1) # people with surg type 1 first but no additional

u_pun <- filter(u,record_id %in% type1firstsurgppl & surgery_number=="1"& punctures!="" & punctures!="Unknown")
pun1 <- unique(filter(u_pun,punctures=="1")$record_id)

# pun1 - 1 punct vs everything else (including unknowns / NAs)
ppl1didx <- unique(intersect(ppl1,pun1)) # surg 1 type 1 and multi surg, had puncture = 1 for their 1st surg
ppl2didx <- unique(intersect(ppl2,pun1)) # surg 1 type 1 and had only 1 surg, had puncture = 1 for their 1st surg

df <- data.frame("Pun 1" = c(length(ppl2didx), length(ppl1didx)), "Not pun1" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df,alternative="less") # was this enriched relative to expected

# punmul - multiple punctures vs everything else (including unknowns / NAs)
punmul <- unique(filter(u_pun,punctures!="1")$record_id) # had multi punctures
ppl1didx <- unique(intersect(ppl1,punmul)) # surg 1 type 1 and multi surg, had multi punctures for their 1st surg
ppl2didx <- unique(intersect(ppl2,punmul)) # surg 1 type 1 and had only surg, had multi punctures for their 1st surg

df <- data.frame("Pun Multi" = c(length(ppl2didx), length(ppl1didx)), "Not Pun Multi" = c(length(ppl2) - length(ppl2didx), length(ppl1) - length(ppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df) # was this enriched relative to expected


# pun1 - comparing 1 puncture vs multi when only considering those that had reported (aka not unknown/NA)
oppl1didx <- unique(intersect(ppl1,pun1))
oppl2didx <- unique(intersect(ppl2,pun1))
mppl1didx <- unique(intersect(ppl1,punmul))
mppl2didx <- unique(intersect(ppl2,punmul))

df <- data.frame("Pun 1" = c(length(oppl2didx), length(oppl1didx)), "Pun Multi" = c(length(mppl2didx), length(mppl1didx)), row.names = c("Surg 1 Type 1 had 1 surg", "Surg 1 Type 1 had > 1 surg"))
df
fisher.test(df, alternative="less") # was this enriched relative to expected

```

```{r logit}

s <- filter(u,MainRow=="Y"&surg1type1=="Y")
s$electrode <- ifelse(s$record_id %in% electrode,"Y","N")

# For the surg 1 type 1 people, TUIdef ~ age at first surg
model <- glm(TUIdefnum ~ agesurg1,data=s,family=binomial)
summary(model)
exp(coef(model))[["agesurg1"]] # odds ratio
exp(confint(model))["agesurg1",] # 95% confidence interval, based on profile likelihood
summary(model)$coef[,"Pr(>|z|)"][["agesurg1"]] # pvalue

model <- glm(TUIdefnum ~ agesurg1 + sex + ureterocele_pole + treatment_plan + PrenatDiagnosis + diagnosis_reason___2+ prereflux + electrode,data=s,family=binomial)
summary(model)
model <- glm(TUIdefnum ~ ureterocele_pole + PrenatDiagnosis + electrode,data=s,family=binomial)
summary(model)
model <- glm(TUIdefnum ~ agesurg1 + PrenatDiagnosis,data=s,family=binomial)
summary(model)

```

